Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2015-06-19 analysis/8-sim-extinction-risk.Rmd
Continue developing projections code

```{r}
library("skewt")
sim_gomp_extinction <- function(abundance_ts, lambda = 1, b = 0.5, N = 200,
  gamma = 0.75, df = 3, thresh = 3, make_plot = FALSE, sigma = 0.2,
  return_dat = FALSE, id = 1) {

  y1 <- log(abundance_ts[length(abundance_ts)])

  proc_error <- rskt(N, df = df, gamma = gamma) / (1 / sigma)
  y <- gompertz(N, lambda, b, proc_error, y1)
  if(make_plot) {
    par(mfrow = c(2, 1))
    plot(y, type = "o")
    abline(h = lambda / b, col = "red")
    abline(h = mean(y), col = "blue")
    plot(exp(y), type = "o")
    abline(h = thresh, col = "green", lwd = 2)
  }
    y_return <- c(log(abundance_ts), y)
    generation <- c(seq(-length(abundance_ts) + 1, 0), seq_along(y))
    data.frame(b = b, N = N, gamma = gamma,
      df = df, sigma = sigma, thresh = thresh, lambda = lambda,
      n_ext = sum(exp(y) < thresh), mean_log_abund = mean(y),
      dat = y_return, generation = generation, main_id = id)
}
```

@@

```{r}
ids <- c(6528, 10127, 10007, 20579)
out_ids <- lapply(ids, function(id_i) {
samples <- 80L
posterior_length <- 2000L
posterior_ids <- round(seq(1, posterior_length, length.out = samples))
out <- lapply(posterior_ids, function(i) {
library("rstan")
id <- id_i
x <- readRDS(paste0("../scratch/base/sm-", id, ".rds"))
x <- rstan::extract(x)
gpdd <- readRDS("gpdd-clean.rds") # to get the last abundance
this_dat <- filter(gpdd, main_id == id)

out_gomp <- sim_gomp_extinction(abundance_ts = this_dat$population_untransformed, 
  lambda = x$lambda[i], 
  b = x$b[i], 
  N = 100,
  gamma = 0.75, 
  df = x$nu[i], 
  thresh = 3, 
  make_plot = FALSE, 
  sigma = x$sigma[i], 
  return_dat = FALSE, 
  id = id)
out_gomp$posterior_sample <- i
out_gomp
 })
out_df <- do.call("rbind", out)
out_df
})
out_df2 <- do.call("rbind", out_ids)

out_df2 <- mutate(out_df2, projection = ifelse(generation > 0, TRUE, FALSE))
p <- ggplot(out_df2, 
  aes(generation, exp(dat), colour = projection, group = posterior_sample)) + 
  geom_line(alpha = 0.2) + 
  geom_vline(xintercept = 0, lty = 2) + 
  theme_bw() + ylab("Abundance") + xlab("Year") + 
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
                labels = scales::trans_format("log10", scales::math_format(10^.x))) + 
  annotation_logticks(sides = "l") + 
  guides(colour = guide_legend(override.aes= list(alpha = 1))) +
  facet_wrap(~main_id, scales = "free_y") 
print(p)

>>>>

```{r}
# library("skewt")
sim_gomp_extinction <- function(abundance_ts, lambda = 1, b = 0.5, N = 100L,
  df = 3, thresh = 3, sigma = 0.2, id = 1, bootstrap_residuals = FALSE, 
  model = "heavy") {

  if (is.null(df)) df <- NA

  log_abundance_ts <- log(abundance_ts)
  y1 <- log_abundance_ts[length(abundance_ts)]

  if (bootstrap_residuals) {
    res <- log_abundance_ts[-1] - (lambda + b * log_abundance_ts[-length(abundance_ts)])
    proc_error <- sample(res, size = N, replace = TRUE) # bootstrap the process residuals
  } else {
    if (model == "heavy") {
      # proc_error <- rskt(N, df = df, gamma = gamma) / (1 / sigma)
      proc_error <- rt(N, df = df) * sigma
    } else {
      proc_error <- rnorm(N, sd = sigma)
    }
  }

  y <- gompertz(N, lambda, b, proc_error, y1)

    y_return <- c(log_abundance_ts, y)
    generation <- c(seq(-length(abundance_ts) + 1, 0), seq_along(y))
    data.frame(b = b, N = N,
      df = df, sigma = sigma, thresh = thresh, lambda = lambda,
      n_ext = sum(exp(y) < thresh), mean_log_abund = mean(y),
      dat = y_return, generation = generation, main_id = id)
}
```

@@

```{r}
ids <- c(6528, 10127, 20579)
library("rstan")
gpdd <- readRDS("gpdd-clean.rds") # to get the last abundance
samples <- 2000L
posterior_length <- 2000L
posterior_ids <- round(seq(1, posterior_length, length.out = samples))

out_ids <- lapply(ids, function(id_i) {
  this_dat <- dplyr::filter(gpdd, main_id == id_i)
  message(paste("main_id:", id_i))
  # x <- readRDS(paste0("gomp-base-normal/sm-", id_i, ".rds"))
  x <- readRDS(paste0("../scratch/gomp-base/sm-", id_i, ".rds"))
  x <- rstan::extract(x)

  out <- lapply(posterior_ids, function(i) {
    out_gomp <- sim_gomp_extinction(abundance_ts = this_dat$population_untransformed, 
      lambda = x$lambda[i], b = x$b[i], N = 10L, df = x$nu[i], 
      thresh = 3, sigma = x$sigma[i], id = id_i, bootstrap_residuals = FALSE, model = "heavy")
    out_gomp$posterior_sample <- i
    out_gomp
  })
  out_df <- do.call("rbind", out)
  out_df
})
out_df2 <- do.call("rbind", out_ids)
out_df2 <- mutate(out_df2, projection = ifelse(generation > 0, TRUE, FALSE))

 # out_df2 <- filter(out_df2, exp(dat) < 10e5)
library(ggplot2)
# p <- ggplot(out_df2, 
#   aes(generation, exp(dat), colour = projection, group = posterior_sample)) + 
#   geom_line(alpha = 0.2) + 
#   xlim(-30, 10) +
#   geom_vline(xintercept = 0, lty = 2) + 
#   geom_hline(yintercept = 10, lty = 2) + 
#   theme_bw() + ylab("Abundance") + xlab("Year") + 
#   scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
#     labels = scales::trans_format("log10", scales::math_format(10^.x)),
#     limits = c(1, 1e5)) + 
#   annotation_logticks(sides = "l") + 
#   guides(colour = guide_legend(override.aes= list(alpha = 1))) +
#   facet_wrap(~main_id) 
# print(p)
# ggsave("heavy-tailed-projections.png", width = 10, height = 6)
cvar <- function(x) mean(x[x<=quantile(x, probs = 0.01)])
group_by(out_df2, main_id) %>% summarise(cvar_ = cvar(exp(dat)))

out_df3 <- out_df2 %>% group_by(main_id, generation, projection) %>%
  summarise(
    min_dat = min(dat, na.rm = TRUE),
    q99 = quantile(dat, probs = 0.01, na.rm = TRUE)
    )

p <- ggplot(out_df3, 
  aes(generation, exp(min_dat), colour = projection)) + 
  geom_line() + 
  geom_line(aes(y = exp(q99)), lty = 2) + 
  geom_vline(xintercept = 0, lty = 2) + 
  geom_hline(yintercept = 10, lty = 2) + 
  xlim(-30, 10) +
  theme_bw() + ylab("Abundance") + xlab("Year") + 
  scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x)),
    limits = c(1, 1e4)) + 
  annotation_logticks(sides = "l") + 
  guides(colour = guide_legend(override.aes= list(alpha = 1))) +
  facet_wrap(~main_id) 
p
